Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-12 main.Rmd
changed to pdf book and base_format: asa article + linear calendar algrabra section shrinked further

   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data. Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach. This work provides tools for creating granularities and exploring the associated time series within a tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
bibliography: [bibliography.bib]
preamble: >
  \setlength {\marginparwidth }{2cm}
  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes}
  \def\mod{~\text{mod}~}
output: 
  rticles::asa_article:
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom up. Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.

@@

The grouping operation:
: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer. The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity. More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$,
$$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$

@@

The altering-tick operation:
: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$. The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of $G_1$ in terms of granules of $G_2$. The altering-tick operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each group has $k$ additional granules of $G_2$.

For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days. We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf]. For more variations of calendar algebra operations, see @Ning2002-tf.

>>>>

   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data. Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach. This work provides tools for creating granularities and exploring the associated time series within a tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
bibliography: [bibliography.bib]
preamble: >
  \setlength {\marginparwidth }{2cm}
  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes}
  \def\mod{~\text{mod}~}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom up. Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities. For a coarser granularity $G_2$ and a finer one $G_1$, a _grouping operation_ partitions some granules of $G_1$  to form a granule of $G_2$. For example, $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$. The _altering-tick operation_ generates new full-integer labeled granularity by periodically expanding or shrinking granules of $G_2$ in terms of granules of $G_1$. For example, <!--due to the uneven length of months, we need to use the altering-tick operation to compute months from days. We can --> consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf]. For more variations of calendar algebra operations, see @Ning2002-tf.

@@

<!-- CONTENT CUT -->
<!-- The grouping operation: -->
<!-- : Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer. The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity. More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$, -->
<!-- $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$ -->
<!-- Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$ -->

@@

<!-- CONTENT CUT -->
<!-- The altering-tick operation: -->
<!-- : Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$. The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of $G_1$ in terms of granules of $G_2$. The altering-tick operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each group has $k$ additional granules of $G_2$. -->

<!-- For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days. We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf]. For more variations of calendar algebra operations, see @Ning2002-tf. -->
